Writing Trash Bin

### 2024-01-07 paper.qmd
added in notes from some overview papers

@@

# 2. Defining Uncertainty
Uncertainty visualisation is not made any easier by the fact that the term "uncertainty" lacks a commonly accepted definition in the literature. @Lipshitz1997 even commented that “there are almost as many definitions of uncertainty as there are treatments of the subject”. This mishmash of terminology leads to a large body of work, all claiming to finding the best visualisation or expression of of "uncertainty" but most don't even seem to agree on what uncertainty is.

## 2.1 Motivating Questions
Signal and noise can only be untangled in the presence of a motivating question. While it is seldom mentioned in the uncertainty visualisation literature, but is still important to understand that "uncertainty" does not exist without a motivating question. The entire process of data analysis, from deciding what should be observed as data through to communicating that data in a plot is governed by human decision and the goal of an analysis. At the philosophical level, applied statistics is simply taking real world entities and boiling them down into probabilistic objects, an ontological process that is largely dependent on our goals [@Otsuka2023]. When we move onto data provenance the issue pursists, as what is kept as data and what is tossed away is determined by the motivation of an analysis and what was previously noise can be shown to become signal depending on the resolution of the question [@Meng2014]. After moving onto modelling this issue continues as each research question can be can be categorised as descriptive, predictive, or causal, each of which has its own appropriate statistical methods and motivation agnostic model selection leads to signal devoid of meaning [@Carlin2023]. Even at the final stages of visualisation a lack of understanding of the motivating question make it difficult to untangle what is signal and what is noise, leaving many uncertainty visualisation studies with conflicting results [@Kinkeldey2014]. These cases highlight that the signal that is being conveyed and the noise that hampens that signal cannot be untangled and therefore communicated without a predetermined motivating question. Therefore any discussion of uncertainty must be done within the context of a motivating question as it is exactly what makes the distinction between signal and noise.

## 2.2 Taxonomies
Within the context of a motivating uncertainty can be classified and 
### 2.2.1 Taxonomies of Uncertainty

@@

### 2.2.1 Taxonomies of Visual Uncertainty

# 3.General Attitudes...
## 3.1 Towards uncertainty in general
## 3.2 Towards uncertainty visualisation

# 4 The current landscape of uncertainty visualisation research
In this section we will first go over the current state of uncertainty visualisation research, which is currently a collection of evaluation studies and review papers, and discuss what it is doing well and what could be done better. Following that, we will provide general reccomendations from the literature.
# 4.1 The landscape
In section 2.1 it was mentioned that uncertainty must be defined within a specific motivating question, otherwise it inherently does not make sense. A large difficulty with the uncertainty visualisation evalusation studies is that this rule is not followed. There are a shockingly large number of evaluation studies that seem to pay no attention to the information that is relevant to the question they are asking, and show participants a selection of seemingly random visualisations from a statistical point of view.

@@

Notes from previous sections that were moved here
- The concept of uncertainty being task dependent is *particularly* salient for uncertainty visualisation, and it is repeatedly identified as a problem in previous reviews of the uncertainty visualisation literature [@Kinkeldey2014; @Hullman2016] as well as across many sub domains and applications [@Wallsten1997; @Munzner2009; @Fischhoff2014; @Meng2021; @Amar2005]. The fact that this conclusion is repeatedly reached shows both the importance and the lack of acknowledgement this concept receives.
- @Fischhoff2014 discusses how decisions with different levels of flexibility should have different uncertainty visualisations. If the user has no choice they should simply be made aware of the risk, if they have several options, the user needs to see the uncertainty in each, if they need to come up with the options themselves, they need to see how things work and how the system fails.
- @Amar2005 suggested a taxonomy for information visualisation based on the types of tasks we use visualisations for and suggest 10 "analytical primitives" that we can then map to visualisations.
- Given that uncertainty estimation and communication seem to be largely task depenent, what questions should we be testing? This leads to several unanswered questions. Is our goal in uncertainty communication to acknowledge and convey the inherrent randomness that exists in all statistical calculations and models? Is it to supress signal that could be the result of random variance? Is it to find the best way to convey the mass function of a random variable? A survey of the literature will leave the answers to these questions unclear, however this inherrent issue has not gone unnoticed.
- Statistics is, at its core, the study of uncertainty. Therefore discussing uncertainty visualisation a a separate sub domain to "normal" data visualisation is inherently confusing. What is typically meant by "uncertainty" visualisation is "noise", that is, we want to present some signal cushioned by its natural variance. Unfortunately, this distinction between "signal" and "noise" is entirely goal dependent.
- @Bella2005 found that most participants were ignorant to the fact that error bars are used for both confidence intervals and standard error bars, two wildly different indicators of precision. (**also the paper sherry sent me**)
# 4.2 Key points from the literature

# 5 Great Examples
- best things you have seen and why they are the best
- can't have a best practices because you need experimental results and my point is that the experiments aren't testing the right things

>>>>

- Can discuss each uncertainty in isolation using distributions, ranges, and lists of alternatives, but the total cumulative uncertainty from every stage is more difficult (near impossible) @Spiegelhalter2017
- "cannot assess the quality of risk communication unless the objectives are clear" circles back to motivation. He assumes we are fulfilling a duty to inform. @Spiegelhalter2017
- the Affect heuristic comes to play in how people understand risk @Spiegelhalter2017
- avoids the topic of risk at times of accute crisis because numerical risk would be inappropriate @Spiegelhalter2017
- positive vs negative framing matters, can try to avoid it @Spiegelhalter2017
- no preference between percent or frequency although (small interaction with numeracy skills though) @Spiegelhalter2017
- Need careful specification of the reference class (e.g. % chance of rain should be fractions of situations like this in which is rains - over both time and location) @Spiegelhalter2017
- Need to make time interval clear (50% each year) @Spiegelhalter2017
- understanding better in a table than text @Spiegelhalter2017
- if expressing as a frequency, note that larger numerators suggest a larger risk, exteme version of this is ignoring the denominator all together @Spiegelhalter2017
- Graphical features that improve accuracy, features that facilitate behavioural change, and features that viewers like may be different (are not necessarily the same). @Spiegelhalter2017
- also should consider graphical literacy @Spiegelhalter2017
- graphs are good for gists and different graphics are good at different things (i.e. trends vs comparison etc) but hard to avoid framing (show information as part of a whole) @Spiegelhalter2017
- scatting in pictograph increases the impression of randomness but makes it difficult to count @Spiegelhalter2017
- people with high numeracy count pictographs but low numeracy assess area @Spiegelhalter2017
- simplifying a multui-outcome array to two outcomes improved understanding @Spiegelhalter2017
- using words to communicate numerical risk is generally discouraged and using words without numerical interpretation can lead to lower accuracy and higher risk aversion @Spiegelhalter2017
- heath had a wide range of different metrics they adopted to communicate chronic risk of adverse events in the future @Spiegelhalter2017
- hard to communicate small risks because people struggle to discrimintate between 1 in 10 000 and 1 in 100 000. Common solution is to use a logarithmic scale (if drawn vertically these are known as risk ladders). design of scale important because risk is associated with position rather than magnitude @Spiegelhalter2017
- compares a bunch of stuff to the chance of getting struck by lightning since they are on a common scale (which is kind of funny) @Spiegelhalter2017
- hierarchy for level of precision for epistemic uncertainty: 1) numbers given to appropriate levels of precision 2) a distribution range 3) a measure of statistical significance 4) verbal quantifiers to numbers 5) refusal to give a number unless the evidence is good enough @Spiegelhalter2017
- Approaches to communicating limited understanding include 1) explicit model uncertainty 2) Qualitative scales expressed as strength of evidence (e.g. high to low certainty of evidence) 3) Acknowledge limitations and their possible impact 4) Acknowledged ignorance @Spiegelhalter2017
- He then went through and explained how uncertainty is communicated across a wide range of areas @Spiegelhalter2017
- Hurricane risk is not only the path, but also the storm surge and wind speed. @Spiegelhalter2017
- coloured maps for earthquake risk are easily interpretable @Spiegelhalter2017
- low-probability high-impact events are particularly difficult to communicate risks about @Spiegelhalter2017
- industrial accidents have a "tollerable" risk chart based on risk and number of fatalities @Spiegelhalter2017

- General issues when communicating risks: clear objectives, segment audience into target groups, use plain language and limit info to only what is necessary, allow for different levels of interest knowledge and numeracy (a gist level, numerical level, then evidence and uncertainty) @Spiegelhalter2017
- Numerical risks: Use absolute risks (but relative risk when catastrophic event), for single unique event use percent chance, express change as a proportion, frequency, or percentage (and be clear about reference class), avoid framing bias by showing risk with and without the outcome, keep the denominator fixed when making comparisons, explicit about time interval, For more knowledgeable audiences consider providing epistemic uncertainty and confidence in analysis @Spiegelhalter2017

@@

# 2. Defining Uncertainty
Uncertainty visualisation is not made any easier by the fact that the term "uncertainty" lacks a commonly accepted definition in the literature. @Lipshitz1997 even commented that “there are almost as many definitions of uncertainty as there are treatments of the subject”. This mishmash of terminology leads to a large body of work, all claiming to finding the best visualisation or expression of of "uncertainty" but most don't even seem to agree on what uncertainty is. 
- mentions uncertainty can be variance, precision, accuracy, reliability, or related concepts @Hullman2016

## 2.2 Taxonomies
Papers that mention uncertainty often use a single sentance definition that suits the purpose of the paper, while papers that seek to define uncertainty use a taxonomy. While

Most definitions of uncertainty are done with a taxonomy. Most papers that provide a definition of uncert Within the context of a motivating uncertainty can be classified and

### 2.2.1 Taxonomies of Uncertainty

@@

### 2.2.1 Taxonomies of Visual Uncertainty
  - Most research had focused on developing taxonomies of uncertainty and developing methods to depict uncertainty visually @Kinkeldey2014
- 5 dichotomous categories of uncertainty can be identified (1) explicit/implicit (directly mapping or showing multiple outcomes) (2) intrinsic/extrinsic (using existing symbols e.g. colour value, or new objects e.g. grids) (3) visually integral/separable (can or cannot be separated from the data) (4) coincidence/adjacent (if data and uncertainty are in integrated or separate views (5) static/dynamic (animation/interaction) @Kinkeldey2014
- Based on those they only use (4), (2), and (5), left out (1) because most visualisations are explicit, and (3) corresponds to (1) in most cases. @Kinkeldey2014
- They sugest that future typologies should take different categoeis of tasks into account (1) communication tasks (2) analytical tasks (3) exploratory takss. List task-centred typology as a main topic for future research in the field to facilitate practical guidelines in this field @Kinkeldey2014

- @Spiegelhalter2017 Establishes that risk is known probabilities and uncertainty is estimated probabilities (i.e. more real world). Suggest uncertainty is what we have when you cannot accurately quantify risk
- Highlights the lack of a clear definition of uncertainty "it depends on who you ask" @Spiegelhalter2017
- ambiguity is also poorly defined, as it can be basically the same as uncertainty (uncertain probabilities) or contested values @Spiegelhalter2017
- A lot of taxonomies for uncertainty, simple is "known knowns, known unknowns, and unknown unknowns", ka (known, unknown, nad unknowable), (risk, uncertainty, ignorance). This also breaks down into the framework of aleatory uncertainty (unavoidable randomness), epistemic uncertatiny (uncertainty about structure, still quantified), ontological uncertainty (uncertain about entire modelling process). @Spiegelhalter2017
# 3.General Attitudes...
I dont think this is worth a section but it is mentioned a lot in .
- the role of unertainty visualisation is to communicate accurate information
- problem of communicating to a group thing (avoid communicating inaccurate information)
- how people feel about uncertainty is irrelevant
  - natural disasters make people uncomfortable but to include it in a conversation about whether or not we should express it is rediculous.

## 3.1 Towards uncertainty in general
## 3.2 Towards uncertainty visualisation

# 4 LITERATURE SUMMARY
## intro
In this section we will first go over the current state of uncertainty visualisation research, which is currently a collection of evaluation studies and review papers, and discuss what it is doing well and what could be done better. Following that, we will provide general reccomendations from the literature.

- methods used in uncertainty visualisation evaluation remain ad hoc. studies approach issue from a usability perspective instead of asking WHY representations DO or DO NOT work. Studies do not follow any methodology commonly agreed upon. @Kinkeldey2014
- Lack of formalisation and eigour in empirical methdos is an issue that is much broader than uncertainty visualisation (extends to info-vis and related domains). Therefore this paper is a starting point, not a definitive summary @Kinkeldey2014
- The question in these papers is often "does method a work better than method B" which is an engineering approach, where the toal is to improve a particular product rather than create general principals @Kinkeldey2014
- Authors then try to generalise beyond the specific constraints of the test even though the framing does not allow for it @Kinkeldey2014
- Alternatively, we should ask how/why does method A work better than method B, which grounds work in perceptual and cognative theory and provides a framwork @Kinkeldey2014

# 4.?? Noise vs Signal (question +)
The first issue that arises with uncertainty research is that uncertainty is frequently treated as a signal, not noise. This discrepancy is also noted by @Kinkeldey2014 who, in their own literature review on uncertainty methods, comment that the literature seems unsure if uncertainty is simply another variable or if it should be treated differently. This is a fair question to ask. If uncertainty is simply a signal we are trying to convey, why should it be treated any different to any other signal in a visualisation?

- The literature also does not discuss if uncertainty is another variable, or if it needs to be treated differently (i,.e. is it metadata of another variable). Most papers studies on uncertainty dont consider this issue, and the question remains, should someone have two separate values, or an integrated view uncertain data view @Kinkeldey2014

This is best seen with a simple look at some papers
(Make into table)
Questions where uncertainty is a signal

Questions where uncertainty is a noise

## 4.?? Greater Infoviz Issues
- Since there are so many papers that treat uncertainty as a signal, not a noise, issues in the uncertainty visualisation literature also help to understand some difficulties in the information visualisation literature as large.

- Suggests that it is straightforward to show a value but much more complex to show uncertainty @Hullman2016 (I think this is interesting in relation to how people SEE uncertainty visualisation. Why is uncertainty hard to visualise?)
- Authors provide little justification for their chosen response models (e.g. absolute accuracy vs relative measures) @Hullman2016
- When asking questions, researchers should focus on events that can be repeated (because of a misunderstanding of confidence intervals) @Hullman2016
- Uncertainty defined as a by product of the process of making a visualisation, from data collection to commucating @Hullman2016
- Some questions (e.g. what are the chances that the no.6 bus will arrive first) you can elicit the viewers subjective probability distribution @Hullman2016
# 4.?? Tasks and Motivating Questions
## 4.1?? In other statistics
What is meant by "uncertainty" may seem obvious to some, but when you attempt to quantify or visualise it you will quickly find yourself asking, "uncertainty about... what?". Do you mean uncertainty on an estimate? On a forecast? How many steps ahead is this forecast? Are we only considering the uncertainty in the estimate or in the parameters or are we considering the possibility of measurement error or biased inputs? Signal and noise can only be untangled in the presence of a motivating question.

The idea that uncertainty can only be defined in the presence of a motivating question is well grounded in most areas of statistics. The entire process of data analysis, from deciding what should be observed as data through to communicating that data in a plot is governed by human decision and the goal of an analysis. At the philosophical level, applied statistics is simply taking real world entities and boiling them down into probabilistic objects, an ontological process that is largely dependent on our goals [@Otsuka2023]. When we move onto data provenance the issue persists, as what is kept as data and what is tossed away is determined by the motivation of an analysis and what was previously noise can be shown to become signal depending on the resolution of the question [@Meng2014]. After moving onto modelling this issue continues as each research question can be can be categorised as descriptive, predictive, or causal, each of which has its own appropriate statistical methods and motivation agnostic model selection leads to signal devoid of meaning [@Carlin2023]. Even at the final stages of visualisation a lack of understanding of the motivating question make it difficult to untangle what is signal and what is noise, leaving many uncertainty visualisation studies with conflicting results [@Kinkeldey2014]. These cases highlight that uncertainty is defined at *every* stage in relation to our motivating question, from data collection to visualisation. Discussions of uncertainty cannot be had if we are not clear *what* we are uncertain about. Once it is established what we are uncertainty about, we can consider the other elements of uncertainty that need to be defined.

## 4.1?? General task notes
- Mentioned alleatory vs epistemic uncertainty, in the visualisation literature alleatory uncertainty is more commonly the focus. also mentions dependence and joint pdfs as a consideration @Hullman2016
- Most studies involve intrinsic approaches such as colour, transparency, and those that used extrinsic techniques typically used glyphs, error bars, grid based techniques, or contouring @Kinkeldey2014
- most studies used coincident approaches, static visualisations, are applied to a specific domain and dont have generalised results @Kinkeldey2014
- The tasks for participants were most commonly value retrival (some where uncertainty and data value are retrived separately, others where they are some separately). also also be asked to extract a relative value @Kinkeldey2014
- How does the complexity of uncertainty relate to the cateogies of user and task @Kinkeldey2014
- Several studies probide evidence that the usability of uncertainty representations can be highly user and task dependent @Kinkeldey2014
- The nature of the task plays an important role for the usability of uncertainty techniques, this may explain many of the insonsitent outcomes from the studies under review since two studies assessing sifferent techniques and user groups are not comparable @Kinkeldey2014
- Doing a follow up publication to discuss issues with reasoning and decision making based on uncertainty visualisations which is not considered or discussed in this paper @Kinkeldey2014
- Main take away is that "we need to systematize future empirical studies on uncertainty visualisation to better enable comparison and generalisation of the findings"
- One way to advance this goal is a taxonomy, however existing taxonomies are focused on data types, uncertainty categories, and representation types @Kinkeldey2014
- he lists a bunch of tasks that were tested with uncertainty visualisationsat the end @Kinkeldey2014

- risk is uncertainty as a signal, and when uncertainty communication is spoken about this is typically what is discussed, uncertainty as signal supression is what is usually meant when we talk about visualisations but that area is functionally ignored.
This idea, that uncertainty can only be defined in the face of a motivating question, is well grounded in the literature but seldom explicitly mentioned

In section 2.1 it was mentioned that uncertainty must be defined within a specific motivating question, otherwise it inherently does not make sense. A large difficulty with the uncertainty visualisation evalusation studies is that this rule is not followed. There are a shockingly large number of evaluation studies that seem to pay no attention to the information that is relevant to the question they are asking, and show participants a selection of seemingly random visualisations from a statistical point of view.

@@

Uncertainty as signal papers have incredibly predictable results
Uncertainty as noise papers should follow a similar protocol to the line up papers

Notes from previous sections that were moved here
- The concept of uncertainty being task dependent is *particularly* salient for uncertainty visualisation, and it is repeatedly identified as a problem in previous reviews of the uncertainty visualisation literature [@Kinkeldey2014; @Hullman2016] as well as across many sub domains and applications [@Wallsten1997; @Munzner2009; @Fischhoff2014; @Meng2021; @Amar2005]. The fact that this conclusion is repeatedly reached shows both the importance and the lack of acknowledgement this concept receives.
- @Fischhoff2014 discusses how decisions with different levels of flexibility should have different uncertainty visualisations. If the user has no choice they should simply be made aware of the risk, if they have several options, the user needs to see the uncertainty in each, if they need to come up with the options themselves, they need to see how things work and how the system fails.
- @Amar2005 suggested a taxonomy for information visualisation based on the types of tasks we use visualisations for and suggest 10 "analytical primitives" that we can then map to visualisations.
- Statistics is, at its core, the study of uncertainty. Therefore discussing uncertainty visualisation a a separate sub domain to "normal" data visualisation is inherently confusing. What is typically meant by "uncertainty" visualisation is "noise", that is, we want to present some signal cushioned by its natural variance. Unfortunately, this distinction between "signal" and "noise" is entirely goal dependent.
- @Bella2005 found that most participants were ignorant to the fact that error bars are used for both confidence intervals and standard error bars, two wildly different indicators of precision. (**also the paper sherry sent me**)

- Need to compare to an approripate ground truth, but this is a philisophical exercse @Hullman2016
## 4.?? Incomparable plots
- But when different techniques are to be compared, it is important that the scenarios and datasets are informationally equivalent, i.e. according to Larkin and Simon (1987, p. 67) ‘[t]wo representations are informationally equivalent if all of the information in the one is also inferable from the other, and vice versa’. A goal in testing, then, is often to determine whether they are also computationally equivalent, or whether one depiction has an advantage over another, i.e. ‘[t]wo representations are computationally equivalent if they are informationally equivalent and, in addition, any inference that can be drawn easily and quickly from the information given explicitly in the one can also be drawn easily and quickly from the information given explicitly in the other, and vice versa’ (Larkin and Simon 1987, p. 67). @Kinkeldey2014

- The way we ask questions and the types of questions we ask are selected with little justification. This paper makes suggestions to reduce the noise in the data from these papers. @Hullman2016

- Most commonly check abolute measures of accuracy, where accuracy is abs(subjective error - actual probabiltiy). Some papers also check relative measures of accuracy and ask subjects to find regions of least certainty or to rank targets by uncertainty. studies also consider response time and confidence rating. some studies also ask users for their own expression of uncertainty (e.g. by asking when sketches are uncertainty and ranking various depictions of uncertainty). Some studies ask subjects to make some decision using the data. ometimes things like complexity of task, degree of visual overload, ease of use, visual appeal, and preferences @Hullman2016
# 4.?? Source of Uncertainty
- rarely used extension is aggregation of uncertainty over an area (retrieve overall estimation from a spatial distribution of uncertainty). Other tasks include comparisons, rankings. also have a task called "search" where participants have to identify entities that fulfill certainty characteristics (high or low values) @Kinkeldey2014
- Whether to use classed or unclassed uncertainty is rarely discussed @Kinkeldey2014. e.g. land map with remotely sensed images has ambiguity in the boundaries, which can result from multiple sources (land cover class, measurement error, images from different dates). If you combine them all it might not make sense for someone who only needs one.

## 4.?? Importance (and unimportance) of Meta information
- Specifies uncertainty can be represented by three components attribute (what) position (where) and temporal (when) @Kinkeldey2014
- Studies dealing with uncertainty information typically deal with attribute but rarely position and time @Kinkeldey2014
- 10/44 studies involve multiple types of uncertainty @Kinkeldey2014

- When we boil down data into a statistical object, we boil away many aspects on context. Birth and coin flips become an object that is one in the same, and often the visualisation 
- Why are spatial and temporal data considered special? What is the statistical element that remains after they have been away? Inexchangeability. Temporal data is one dimensional inexchangeability, spatial data has two dimensional inexchangeability.
  - There are other things that have inexchangeability, and not every question related to spatial or temporal data requires the inexchangeability.
  - 
- Yes the data type is relevant but its not the most important thing
  - one with temporal, one with that temporal paper but you didn't know it was temporal, one with the bus time cdf
  - screenshot from these three papers, the connecting factor is the question, not the data type
  - The data type: spatial, temporal, etc, usually is mentioned as the 
  - Not a feature of the data, but it is a feature of the question "how do I show uncertainty on a map?". This is a fine question to ask, but the lack of specificity in what is being done, and the zoning in on functionally irrelevant elements leads the literature as a huge mess.

# 4.?? Required Changes and Considerations to the Experimental Space 
- If you just keep using rankings, you need to compare everything one by one forever to get an idea of the space @Hullman2016
- people sometimes ignore uncertainty and instead look at the pattern (e.g. distance between means) and use a heuristic to answer the question. be proactive about possible heuristics, look for signes of heuristics in responses, ask subjects to describe their strategy, consider including degree of heuristic as a dependent variable @Hullman2016
- People believe in the law of small numbers but dont believe in the lar of large numbers @Hullman2016
- experiencing an event makes your percieved probability higher than if you see a description of it @Hullman2016
- Blur and fog are suggested because they are metaphors for uncertainty, however the usefulness of these metaphors is rarely investigated @Kinkeldey2014
- Frequency thing means that more effective communication to participants in a study means more effective science and we have a roll on effect @Hullman2016
- Probability judgements are less relative than other psychophysical judgements, 0%, 50% and 100% all have well understood meanings. Subjective probability distributions are spontaneous, and subjects are constructing a distribution upon being asked rather than articulating a fully formed mental representation of their beliefs. Therefore it is subject to heuristics. consider anchoring a scale with farmiliar probabilities . not sure if being asked for the entire distribution helps, e.g. the graphical one reduces noise but asking people for an interval estimation leads to overprecision and overconfidence @Hullman2016

- If the goal is to compare visualisation, ground truth is easy, if the goal is to evaluate how accurate the subjective probability distributions are, the ground truth becomes complex. ground truth issue e.g. if you provide a sample from a distribution, is the true value the mean of that sample or the population mean? @Hullman2016
- believed that offering enough decisions with varying rewards you can observe the true subjective probability, however whether or not probability is the right representation of subjective uncertainty is a topic of debate @Hullman2016
- she also discusses graphical interface (users have to draw the probability distribution) and framing probabilities as natural frequencies are better choices because they reduce noise and stuff.@Hullman2016
- Consider incorperating utility functions. Include "probability-coherence" checks @Hullman2016
- evidence that people do not follow the axioms of probability when conveying subjective probabilities. How do you deal with this as an analyst (e.g. if the probabilities do not sum to 1 @Hullman2016
- If someone answers a question correctly it is hard to tell if it was because of the information provided or if the person was using a heuristic @Hullman2016
- You cant tell if subjects in an experiment do poorly because the visualisation was bad or because the question was misinterpreted @Hullman2016

- Visualisations risk: consider a good summary table as a visualisation, use multiple formats because no single representation suits all members of an audience, illuminate graphics with words and numbers, design graphics to allow a part-to-whole comparison on an appropriate scale, narrative labels are important (show magnitude through tick marks), use narrative, images and metaphors to gain/retain attention but dont arouse undue emotion, assume low numeracy of a general public audience and adopt a less-is-more approach by reducing the need for inderence and making clear and explicit comparison, be cautious about interactivity and animations (may introduce uncessary complexity), avoid chart junk (like 3d bar charts), assess the needs of the audience, and iterate towards a final design.
# 5 Great Examples
- best things you have seen and why they are the best
- can't have a best practices because you need experimental results and my point is that the experiments aren't testing the right things
- VSUP, the map interactive map that susuan showed me, the pixel map

### 2024-01-07 writingdumpster.qmd
added in notes from some overview papers

# context section outtakes
This conceptual way of thinking about uncertainty has been rattling around in the different areas of data science for a while.
For example, when we record coin flips we typically want to model the behavior with a binomial distribution, so we ignore the outcome of a coin landing on its side despite the fact it is a real world possibility.
@Wallsten1997 argue that the best method for evaluating or combining subjective probabilities depends on the uncertainty the decision maker wants to represent and why it matters.  @Munzner2009 created a nested model for visualisation that highlighted how pitfalls in upstream levels, such as problem characterisation and data abstraction, can cause downstream effects and damage the effectivness of a visualisation. @Fischhoff2014 looks at uncertainty visualisation for decision making decides that we should have different ways of communicating uncertainty based off what the user is supposed to do with it.

>>>>

# context section outtakes
Given that uncertainty estimation and communication seem to be largely task depenent, what questions should we be testing? This leads to several unanswered questions. Is our goal in uncertainty communication to acknowledge and convey the inherrent randomness that exists in all statistical calculations and models? Is it to supress signal that could be the result of random variance? Is it to find the best way to convey the mass function of a random variable? A survey of the literature will leave the answers to these questions unclear, however this inherrent issue has not gone unnoticed.
This conceptual way of thinking about uncertainty has been rattling around in the different areas of data science for a while.
For example, when we record coin flips we typically want to model the behavior with a binomial distribution, so we ignore the outcome of a coin landing on its side despite the fact it is a real world possibility.
@Wallsten1997 argue that the best method for evaluating or combining subjective probabilities depends on the uncertainty the decision maker wants to represent and why it matters.  @Munzner2009 created a nested model for visualisation that highlighted how pitfalls in upstream levels, such as problem characterisation and data abstraction, can cause downstream effects and damage the effectivness of a visualisation. @Fischhoff2014 looks at uncertainty visualisation for decision making decides that we should have different ways of communicating uncertainty based off what the user is supposed to do with it.
